---
title: "Lab 1 - Data visualization"
format: html
editor: visual
author: "Prabhjot Singh"
---

# Questions

## Part 1

### Question 1

```{r}
library(tidyverse)
library(ggplot2)
library(plotly)
```

#### Midwest Dataset

```{r}
midwest
```

```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100"
  )
```

```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 1000) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 1000"
  )
```

```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 10000) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 10000"
  )
```

```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100000) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100000"
  )
```
#### Comment: Histogram with binwidth set to 1000 is the most appropriate as it makes it easy to infer details about the data on the other hand, when binwidth is set to 100 it is hard to make out anything as everything is very close to each and when set to 10000 or 100000, the bins are too broad to infer anything.

### Question 2
```{r}
ggplot(midwest,aes(x = popdensity))+geom_density()+labs(x="Population Density",y="Count",title="Population Density of Midwestern Counties")
```
```{r}
midwest|>select(county, popdensity)|>arrange(desc(popdensity))
midwest|>summarise(population_density=mean(popdensity),n=n())

```
#### Comment
The overall range of population density is very wide and right-skewed. Most counties have an average population density of about 3,097 people per unit area. Across the full range, most counties fall on the lower end of the scale, with only a handful having significantly higher population densities. One clear outlier is Cook County, with a population density of 88,018.40, which makes sense given that it is the most populous county in the U.S.

### Question 3
```{r}
p<-ggplot(midwest,aes(x=percollege,y=percbelowpoverty,colour = state,shape = state))+geom_point()+labs(x="Percentage with College Degree",y="Percentage Below Poverty",title="Percentage of People with College Degree vs Below Poverty",subtitle = "By State")
p
ggplotly(p)
```
#### Comment
There is a downward trend between the percentage of people with a college degree and the percentage of people below the poverty line, which means that, generally, as people become more educated, they rise above the poverty line. There is one noticeable outlier: Menominee County in Wisconsin, where 7.3% of the population has a college degree and 48.7% are below the poverty line. In most counties, the general trend of decay holds, with some dissimilarities, such as in Ohio, where counties fall within a similar range, and in Illinois, which has a significant outlier: Jackson County, where 36.6% of people have a college degree and 28.4% are below the poverty line.

### Question 4
```{r}
p<-ggplot(midwest,aes(x=percollege,y=percbelowpoverty,colour = state,shape = state))+geom_point()+geom_smooth(se=FALSE)+facet_wrap(~state)+labs(x="Percentage with College Degree",y="Percentage Below Poverty",title="Percentage of People with College Degree vs Below Poverty",subtitle = "By State")
p

```

#### Comment
I prefer this graph over the previous one, as the correlations between the variables can be easily identified. The addition of the smooth curve greatly helps in establishing outliers and highlighting the regions where most points are concentrated.

### Question 5
```{r}
p<-ggplot(midwest,aes(x=area,y=state,colour = state))+geom_boxplot()+labs(x="Area of County (Units Unknown) ",y="State",title="Area of County by State")
p

```
#### Comment
While the range of county area sizes is consistent across states, Wisconsin has the largest interquartile range, with counties that include both the smallest area (area = 0.014) and one of the largest areas (area = 0.094) without being considered outliers. In contrast, Indiana has the smallest interquartile range. The most extreme outlier is Marquette County in Michigan.

### Question 6
```{r}
#Creating another variable based on the inmetro value, if 1 then yes otherwise no
midwest <- midwest |>
  mutate(metro = if_else(inmetro == 1, "Yes", "No"))
p<-ggplot(midwest,aes(x=state,fill = metro))+labs(x="State",y="Proportion",title="Metro vs Non-Metro Counties",subtitle = "by State")+geom_bar()+scale_y_continuous(labels=scales::label_percent(suffix = "",scale = 1/100))
p


```

#### Comment
Generally, in all of the given states, most counties are not considered a metro area. In Wisconsin, about 74% of the counties are classified as non-metro and about 25% as metro. Ohio has a more balanced distribution, with roughly half(50%) of its counties considered metro and the other half non-metro. Illinois(), Indiana, and Michigan each have a greater number of counties classified as non-metro than metro.

### Question 7
```{r}
p<-ggplot(midwest,aes(x=percollege,y=percbelowpoverty,colour = percwhite))+labs(x="% College Educated",y="% Below the Poverty Line",title="Percentage of Population with College Degree vs Below Poverty Line",subtitle = "with racial composition per county")+geom_point()+facet_wrap(~state)+theme()
p
ggplotly(p)
```

#### Comment
Menominee County is the greatest outlier in Wisconsin.
When compared to other racial compositions (Black Americans, Whites, Asians and other races) American Indians have the highest percentage in Menominee and with almost no Black Americans there.
## Part 2

**Enough about the Midwest!**
```{r}
nc_county <- read_csv("data/nc-county.csv")
```

### Question 8
#### Comment
Usually, there is a positive correlation between Land Area and Population Density as people like spacious and bigger places to live and raise their families.
```{r}
p<-ggplot(nc_county,aes(x=land_area_mi2,y=density))+labs(x="Land Area (Square Miles)",y="Population Density",title="Land Area vs Population Density")+geom_point()+geom_smooth(se =  FALSE)
p

```
#### Comment
There is an almost straight, or stable, relationship between land area and population density. It looks like my guess was incorrect. Apparently, there is no positive or negative relationship between the variables.

### Question 9
```{r}
p<-ggplot(nc_county,aes(x=land_area_m2,y=density))+labs(x="Land Area (Square Meters)",y="Population Density",title="Land Area vs Population Density")+geom_point()+geom_smooth(se =  FALSE)
p

```
#### Comment
This graph is very similar to the one in the last question, showing the same trend. This is because the unit of measurement doesn’t affect the relationship between the variables.